Adaptive Scheduling of Scientific Workflows

### 2024-04-16 main.tex
described vanilla HEFT without mems, corrected the description of heuristics

    \section{Algorithms}

    \subsection{Baseline: HEFT with memories}

    \algo{HEFT} is one of the most popular scheduling heuristics(TODO cite).
    This algorithm schedules tasks on heterogeneous computing systems with the aim to minimize
    the workflow's overall completion time.
    The classical formulation of \algo{HEFT} does not respect memory boundaries, so we suggest a
    slightly changed version that also takes memory sizes into account.

    \algo{HEFT} first assigns all tasks their ranks on all processors.
    This rank includes the bottom level of the task, including the maximum cost of its successors on the same processor.
    Then, for each task, they are being assigned to the processor with the best rank.
    The algorithm looks to minimize the earliest finish time of the task when it chooses the processor with the best rank.
    Finally, the algorithm calculates the order of execution of tasks on each processor, depending on their dependencies.

    We only change the assignment to processor part.
    Before assigning a new task there, we calculate the memory requirement of already assigned tasks.
    If the remaining memory is big enough to host the task, we assign it there, otherwise we try the next-best processor
    rank-wise.

    \skug{Even easier: calculate the memory requirement of already assigned tasks by summing them up. This will lead
    to an overestimation of memory requirement.}

@@

    \subsection{Limitations}

    We assume that moving a task from one processor to another costs nothing, when in reality it requires a copy
    operation that can take time at least equal to the sum of the weights of all its input files divided by the bandwidth.

We also keep track of the current memory usage, $\curM_j$ and $\curC_j$,
on respectively the processor memory and communication buffer. 
Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
They correspond to some $c_{u,v}$'s.

\smallskip
\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
\[Res = \curM_j  - m_v + \sum_{u \in Pred(v), u\notin T(p_j)}  \{c_{u,v}\}
- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]

\smallskip
\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
    Its starting time $st_v$
is dictated by the maximum between $rt_j$, and all communications that
must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is then
    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
    Its finish time on $p_j$ is then
$FT(v,p_j) = st_v + \frac{w_v}{s_j}$.

\medskip
\noindent{\em Assignment of task~$v$.}\\
Once we have computed all finish times for task~$v$, 
we keep the minimum $FT(v,p_j)$ and assign task~$v$
to processor~$p_j$.
In detail, we:

    \textbullet
    Evict the memories that need to be evicted to free the memory. We remove these memories from pending memories
        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.

    \textbullet For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
    communication time $c( u,v) / \beta$.

    \textbullet We compute the correct amount of available memory for $p_j$ (for when the task is done).
For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
    the incoming edge, also freeing the same amount of available memory (increasing $CurM_j$).
For each successor, on the other hand, we add the edge weights to pending memories and reduce $CurM_j$ by the corresponding
    amount.

    updating $\MM_j$ and $\MC_j$ as needed,
as well as all the $rt$ values for $p_j$ and the communication channels.

>>>>

We also keep track of the currently available memory, $availM_j$ and $availC_j$,
on respectively the processor memory and communication buffer. 
Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
They correspond to some $c_{u,v}$'s.

\subsection{Baseline: original HEFT without memories}

    Original HEFT does not consider memory sizes.
    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.

    HEFT works in two steps.
    In the first step, it calculates the ranks of tasks by computing their non-increasing bottom levels.
    The bottom level of a task is defined as
    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
    (the max is 0 if there is no outgoing edge).
    The tasks are sorted by non-decreasing ranks.

    In the second step, the algorithm iterates over the ranks and tries to assign the task to the processor where it
    has the earliest finish time.
    We tentatively assign each task to each processor.
    The task's starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is then
    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
    Its finish time on $p_j$ is then
    $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.

    Once we have computed all finish times for task~$v$,
    we keep the minimum $FT(v,p_j)$ and assign task~$v$
    to processor~$p_j$.

    \textit{Assignment to processor}
    When assigning the task, we set the ready time of the processor~$j$ $rt_j$ to the finish time of the task.
    For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
    communication time $c( u,v) / \beta$.

\smallskip
\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
\[Res = availM_j  - m_v - \sum_{u \in \Pi(v), u\notin T(p_j)}  \{c_{u,v}\}
- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]

\smallskip
\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
    Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is then
    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} } \]
    Its finish time on $p_j$ is then
$FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.

\medskip
\noindent{\em Assignment of task~$v$.}\\
Once we have computed all finish times for task~$v$, 
we keep the minimum $FT(v,p_j)$ and assign task~$v$
to processor~$p_j$.
In detail, we:
    \begin{itemize}
        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
        We remove these files from pending memories
        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
        \item    Calculate the new $availM_j$ on the processor.
        We subtract the weights of all incoming files from predecessors assigned to the same processor,
        and add the weights of outgoing files generated by the currently assigned task.
        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
        communication time $c( u,v) / \beta$.
        We also remove the incoming files from either the pending memories or pending data in buffers of these other
        processors, and increase the available memories or available buffer sizes on these processors.
        \item We compute the correct amount of available memory for $p_j$ (for when the task is done).
        For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
        the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
        For each successor, on the other hand, we add the edge weights to pending memories and reduce $availM_j$ by the corresponding
        amount.
    \end{itemize}

\section{Appendix}

@@
